---
title: "SkinCell_Analysis"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
 
Skincells dataset has 3 variables, day of observation, amount of time colony was exposed to solar simulator, number of live cells  in the colony. Experiment was performed on 4 days, and collected 118 observations.

```{r,echo=FALSE} 
# Load the dataset into RStudio 
data <- read.csv("temp.txt",sep = '\t')
```

Sample data can be seen in Table1.

```{r}
head(data)
```

## Descriptive Analysis

Descriptive analysis is performed on the data and below are the observations
1. Number of observations collected on Day1 to Day3 are 29, and Day4 is 31. 
2. Time exposed in mins, ranges from 0 to 4 minutes
3. logcells values range from 0 to 12 in count. At first glance, it is evident that the longer time the cells are exposed, the shorter their life span is.

Refer Figure-1 below

```{r, echo=FALSE}
library(ggplot2)
g <- ggplot(data, aes(factor(time), logcells)) + labs(x="LogCells",y="Time in mins") + ggtitle("Figure-1 Cell Count decreases with Increased Exposure")
```

```{r}
g   + geom_boxplot(outlier.colour = "green")  + coord_flip()
```
 
A linear regression model was fitted to these data with logcells as a continuous response variable,  time is a continous predictor, day is a categorical predictor. To allow maximum flexibility in the initial model, an interaction of day and time was included.


## Model - 1 

```{r,echo=TRUE}
data$day <- as.factor(data$day)
reg.model<-lm(logcells~(time+day+day*time),data=data)
summary(reg.model)
anova(reg.model)
```

p-value is statistically significant for time, indicating that cells count is predominantly dependent on exposure time.
Every unit of increased exposure time, cells count decreases by 1.4 units
Prediction accuracy of the model is 26%, which is very low.
Outlier detection analysis is performed to check, if there were incorrect observations captured due to contamination or any other factors.

```{r,echo=FALSE}
library("car")
# Outlier Analysis 
influence.measures(reg.model)
influenceIndexPlot(reg.model, id.n=3) # Index Plots of the influence measures
```

```{r}
# Outlier Analysis : 
influencePlot(reg.model, id.n=3) # A user friendly representation of the above
```

Below observations are severly influencing the model. 

```{r}
data[c(20,28,29,56,86,88,116),]
```

Data points 28, 86 are contradicting to our understanding, i.e When the exposure is maximum, cells count should be low.
however in Data point 86, cell count is 9. This could have caused due to contamination or incorrect results capture.

## Model - 2

```{r}
# Deleting 20,28,29,56,86,88,116 data points 
reg.model2<-lm(logcells~(time+day+day*time),data=data[-c(20,28,29,56,86,88,116),])
summary(reg.model2)
```

Prediction accuracy increased to 32% after removal of outlier datapoints.
Many of the interaction variables are insignificant, and their importance can be checked by performing stepwise model selectoin by AIC

```{r}
# Perform stepwise Model selection
library("MASS")
stepAIC(reg.model2)
```

AIC is lowest, when the model includes time and day variable, without interaction. Lets build the final model and check the prediction accuracy of the model.

## Model - 3

```{r}
# Deleting 20,28,29,56,86,88,116 data points and day and time are the only predictors
reg.model3<-lm(logcells~(time+day),data=data[-c(20,28,29,56,86,88,116),])
summary(reg.model3)
```

Prediction accuracy jumped to 34%. 

# Conclusion

1. Every unit increase of exposure time, reduces cells count by 1.3
2. day2, day3, day4 are having significant influence on cell count, which indicates there is definitely an external influence which impacted the experiment.
3. It is advisable to perform the experiment in a controlled environment for better results and predictability.
